Embedding Dimension of Polyhedral Losses

### 2020-01-30 colt-20/embed-dim.tex
Removing ^r from the polytope sets

\paragraph{Embeddings.}
%We specifically study polyhedral functions in part because they have high-dimensional subgradient sets at points of nondifferentiability, which will be of interest to us later.
Polyhedral losses are motivated partly because they correspond to a natural surrogate construction technique: embedding the discrete predictions $\R$ as points in $\reals^d$ \jessie{and linearly interpolating between embedded reports}.
The key condition required on the embedding and surrogate loss $L$ is that a discrete prediction $r \in \R$ should minimize expected $\ell$-loss if and only if its embedding point minimizes expected $L$-loss.
%
\begin{definition}[Embedding of a loss]
  A loss $L: \reals^d \to \nonnegreals^{\Y}$ \emph{embeds} a loss $\ell: \R \to \nonnegreals^{\Y}$ \emph{in dimension $d$} with \emph{embedding function $\phi: \R \to \reals^d$} if: (i) $\phi$ is injective; (ii) for all $r \in \R$, $\ell(r) = L(\phi(r))$; and (iii) for all $r \in \R$, $\gamma_r = \Gamma_{\phi(r)}$.
  We simply say $L$ \emph{embeds} $\ell$ (in dimension $d$) if some such embedding function exists.
\end{definition}
%
Note that embedding does not immediately imply a key desirable property, consistency, which is equivalent to calibration in finite-outcome settings.
Informally, calibration means that if one minimizes the expected surrogate loss arbitrarily well over $\reals^d$, then applies the link function to obtain a discrete prediction $r$, then $r$ exactly minimizes $\ell$-loss.
\begin{definition}[Calibrated surrogate]\label{def:calibration}
  A surrogate loss $L: \reals^d \to \nonnegreals^{\Y}$ and \emph{link function} $\psi: \reals^d \to \R$ are \emph{calibrated} for a loss $\ell: \R \to \nonnegreals^{\Y}$ if for all $p \in \simplex$,
    \[ \inf_{u \in \reals^d ~:~ \psi(u) \not\in \gamma(p)} \inprod{p}{L(u)}  > \inf_{u \in \reals^d} \inprod{p}{L(u)}  .\]
  We simply say $L$ can be \emph{calibrated} to $\ell$ if such a link function exists.
\end{definition}
\bo{More on why calibration is desirable?}
\jessie{Taking a crack at this:}
As we alluded to above, calibration is not actually the trait we desire of a surrogate; rather, we seek consistency.
\cite{tewari2007consistency} show that, in a finite-outcome setting, the presence of calibration and consistency are equivalent conditions.
By consistency, we mean the standard notion where the expected loss over $\X \times \Y$ for a sequence of hypotheses $h_m$ converging to the optimal expected surrogate loss implies the expected discrete loss over the linked hypotheses $\psi \circ h_m$ converges to the optimal discrete loss.
The equivalence of these conditions allows us to focus on calibration, so we can abstract away the input space $\X$ and simply focus on probability distributions over the outcomes $\Y$.

@@

Therefore, we abstract to consider collections of sets $T^r_y$, which aspire to be the subgradient sets of a calibrated polyhedral surrogate $L(\cdot)_y$ at $u_r$.
Note that if $L(\cdot)_y$ is a polyhedral function on $\reals^d$, then all of its subgradient sets are (bounded) closed polytopes.\bo{cite}\jessie{\cite{rockafellar1997convex} ? in my notes, but I don't have the exact theorem number.}

\begin{definition}[$\T$, $\T^r$, $D(\T^r)$]
  We use the following notation for a collection of closed polytopes, with implicit parameter $d$: $\T = \{T^r_y \subseteq \reals^d : r \in \R, y \in \Y\}$.
  For each fixed $r$, we write $\T^r := \{T^r_y : y \in \Y\}$.

  Given a distribution $p \in \simplex$, we write the $p$-weighted Minkowski sum of $\T^r$ as
    \[ \oplus_p \T^r := \left\{ \sum_{y \in \Y} p(y) x^r_y ~\Big|~ x^r_1 \in T^r_1, \dots, x^r_y \in T^r_y \right\} , \]
  in other words, the Minkowski sum of the scaled sets $\{p(y) T^r_y : y \in \Y\}$.

  Finally, we associate with $\T^r$ a set of distributions $D(\T^r) = \{ p \in \simplex : \vec 0 \in \oplus_p \T^r\}$.
\end{definition}

The importance of the $p$-weighted Minkowski sum and of $D(\T^r)$ are that they capture the distributions $p$ for which $u_r$ minimizes expected loss, assuming that $\T$ correspond to some polyhedral $L$.
\begin{proposition} \label{prop:minkowski-0-opt}
  If $T^r_y$ is the subgradient set of each $L(u_r)_y$, then $D(\T^r)$ is the set of distributions $p \in \simplex$ for which the embedding point $u_r$ minimizes expected $L$-loss.
\end{proposition}
\begin{proof}
  Any convex function $f$ is minimized at $u$ if and only if $\vec 0$ is a subgradient of $f$ at $u$.
  Let $f(u) = \inprod{p}{L(u)}$.
  Then the subgradient set of $f$ at $u$ is $\oplus_p \T^r$.
  This follows from the basic properties that if $f_1,f_2$ are convex with subgradient sets $T_1,T_2$ at $u$, then $\alpha f_1$ has subgradient set $\alpha T_1$ and $f_1 + f_2$ has subgradient set $T_1 \oplus T_2$, the Minkowski sum.
  So $\vec 0$ is a subgradient of $\inprod{p}{L(u_r)}$ if and only if $\vec 0 \in \oplus_p \T^r$, which by definition occurs if and only if $p \in D(\T^r)$.
\end{proof}
This fact will be vital for characterizing when $\ell$ is correctly embedded by some $L$ whose subgradient sets are $\T$.
We leverage it next.

@@

Two conditions are required: \emph{optimality} and \emph{monotonicity}.
Optimality enforces that the surrogate is minimized precisely when and where it should be.
It says that for each discrete prediction $r$ and set of distributions $\gamma_r$ for which it is optimal, there exists a collection of polytopes $\T^r$ (presumably the subgradient sets of $L(u_r)_y$ for each $y$) realizing Proposition \ref{prop:minkowski-0-opt}.
Monotonicity says that these individual polytopes can actually be glued together to form the subgradients of some convex loss function $L$.

@@

\begin{theorem} \label{thm:general-char}
  Let $\ell: \R \to \reals_{\geq 0}^{\Y}$ be a discrete loss with, for each $r \in \R$, $\gamma_r = \{p : r \in \argmin_{r'} \inprod{p}{\ell(r)}\}$.
  Then $\ell$ is $d$-embeddable if and only if there exist polytopes $\T = \{T^r_y \subseteq \reals^d : r \in \R, y \in \Y\}$ such that both of the following hold:
  \begin{enumerate}
    \item (Optimality) For all $r \in \R$, we have $D(\T^r) = \gamma_r$.
    \item (Monotonicity) There exists an injective embedding function $\phi : \R \to \reals^d$ and loss functions $\{L_y : \reals^d \to \reals^\Y_{\geq 0}\}_{y \in \Y}$ such that for all $y \in \Y$ and $r \in \R$, we have $T^r_y = \partial L_y(\phi(r))$.
  \end{enumerate}
\end{theorem}
\begin{proof}
  \jessie{An attempt:}
  $(\implies)$ If $\ell$ is $d$-embeddable, say that $L:\reals^d \to \reals^\Y_{\geq 0}$ embeds $\ell$.
  Take $T^r_y$ to be $\partial L(\phi(r))_y$ for each $r \in \R$ and $y \in \Y$.
  Clearly this set of polytopes $\T$ satisfies the monotonicity condition.
  To see the optimality condition is satisfied, consider for each $r \in \R$ that $\gamma_r = \Gamma_{\phi(r)} = \{p \in \simplex : \vec 0 \in \oplus_p \partial L(\phi(r))_y \} = \{p : \vec 0 \in \oplus_p T^r_y \} = D(\T^r)$.

  \bigskip
  $(\impliedby)$ The first embedding condition holds by the assumption of $\phi$ in the monotonicity condition.
  Now consider that $\gamma_r = D(\T^r) = \{p \in \simplex : \vec 0 \in \oplus_p T^r_y \} = \{p \in \simplex : 0 \in \oplus_p \partial L(\phi(r))_y \} = \{p : r \in \Gamma(p) \} =  \Gamma_{\phi(r)}$.
  Therefore, as this holds for all $r \in \R$, we have that the Bayes Risks of the two losses match, and by~\cite[Proposition 1]{finocchiaro2019embedding}, we have $L$ embedding $\ell$. 
\end{proof}

@@

We now investigate the optimality condition of Theorem \ref{thm:general-char}, for two purposes.
First, we aim to greatly narrow the search space for constructing low-dimensional surrogate loss functions for a given discrete loss.
The tools we construct in this section aid in this task by constraining or constructing feasible subgradient sets $\T^r$ given a level set $\gamma_r$.
Second, we wish to prove impossibilities, i.e. lower bounds on the embedding dimension of a given discrete loss (an apparently hard problem).

By dropping monotonicity from Theorem \ref{thm:general-char}, we obtain $|\R|$ independent optimality conditions, $D(\T^r) = \gamma_r$, that must be satisfied in order to have a $d$-dimensional embedding.
For upper bounds, one must construct a $\T^r$ satisfying $D(\T^r) = \gamma_r$ for all $r$ (and then turn to satisfying monotonicity).
For lower bounds, it suffices to prove nonexistence of any such $\T^r$, for any $r$.

To accomplish these, we construct a quadratic feasibility program (QFP, Definition \ref{def:qfp}) that takes as input a polytope $C \subseteq \simplex$ and a parameter $d \in \mathbb{N}$ and (roughly) attempts to construct $\T^r$.
We will prove the following main results.
\begin{theorem} \label{thm:opt-iff-qfp}
  Given a convex polytope $C \subseteq \simplex$, there exist polytopes $\T^r$ in $\reals^d$ such that $D(\T^r) = C$ if and only if there is a feasible solution to the QFP (Definition \ref{def:qfp}) with parameter $d$.
\end{theorem}

@@

The variables are a set of normal vectors $\{v_i \in \reals^d : i \in [k]\}$ and a set of vertices $\{x^j_y \in \reals^d : j \in [\ell], y \in \Y\}$.
These correspond to a (possibly relaxed) halfspace representation and (possibly incomplete) vertex representation of the subgradient polytopes $\T^r$.

@@

To prove Theorem \ref{thm:opt-iff-qfp}, it is useful to develop a less-constructive intermediate characterization of when $D(\T^r) = C$.
The following condition formalizes a necessary condition on $\T^r$ in terms of the halfspace representation of $C$; the subsequent one formalizes a sufficient condition using the vertex representation.

\begin{condition}[Halfspace condition]\label{cond:H-condition}
	A collection of polytopes $\T^r$ and a polytope $C$ in the simplex defined by $C = \{p \in \simplex : Bp \geq \vec 0\}$ \emph{satisfy the halfspace condition} if there exist $v_1, \ldots, v_k \in \reals^d$ such that, for all $i \in [k]$ and $y \in \Y$, for all $x \in T^r_y$, we have $\inprod{v_i}{x} \leq B_{iy}$.
\end{condition}
\begin{condition}[Vertex condition]\label{cond:V-condition}
	A collection of polytopes $\T^r$ and a polytope $C$ in the simplex defined by $C = \conv(\{p^1, \ldots, p^\ell\})$ \emph{satisfy the vertex condition} if for all $j \in [\ell]$, $0 \in \oplus_{p^j} \T^r$. %% Equivalent: there exist $\{x_{jy} \in T_y : j \in [\ell]\}$ such that, for all $j \in [\ell]$, $\sum_{y \in \Y} p^j_y x_{jy} = \vec 0$.
\end{condition}

To gain a better understanding about the QFP in Definition~\ref{def:qfp}, we prove the following:
\begin{lemma} \label{lemma:D-polytope}
  For any $\T^r$, $D(\T^r)$ is a polytope (in particular, is convex).
\end{lemma}
\begin{proof}
   Recall by definition, the notation $\oplus_p \T^r = \{\sum_y p(y) x^r_y : x^r_y \in T^r_y (\forall y)\}$.
   Each $T^r_y$ is a polytope, so $p(y) T^r_y$ is a polytope.
   The Minkowski sum of polytopes is a polytope, so $\oplus_p \T^r$ is a polytope by~\cite[Section 1.2]{weibel2007minkowski}.
  \bo{TODO}\jessie{Done?}
\end{proof}
\begin{lemma}  \label{lemma:minkowski-support}
  Given polytopes $\T^r$, there exists a finite set of normal vectors $w_1,\ldots,w_K \in \reals^d$ such that, for all $p \in \simplex$, $\oplus_p \T^r = \{x : \inprod{w_i}{x} \leq \sum_{y \in \Y} p(y) \max_{x \in T^r_y} \inprod{w_i}{x} \}$.
\end{lemma}
\begin{proof}
   For each $p$, $\oplus_p \T^r$ is a polytope.
  \bo{TODO: it turns out that $p$ doesn't matter, only its support; and there are only finitely many possible supports (or even better, the normals that are complete for the uniform minkowski sum turn out to be complete for all other $p$-combinations).}
\end{proof}
\begin{lemma} \label{lemma:E-to-B}
  Let $C = \{p : Bp \geq \vec 0 \}$ where $B$ has the minimum possible number of rows to capture $C$, and suppose $C = \{p : Ep \geq \vec 0 \}$.
  Then for each row in $B$ there is some (unique) row in $E$ that is equal to $\alpha B$ for some positive $\alpha$.
\end{lemma}
\begin{proof}
  \bo{? Cite} \jessie{Is left as an exercise in the following text:} \cite[Exercise 2.15]{ziegler2012lectures}
\end{proof}

\begin{theorem} \label{thm:vertex-halfspace-opt}
  Let the polytopes $\T^r = \{T^r_y \subseteq \reals^d : y \in \Y\}$ and $C$ be given, with $C = \conv(\{p^1,\ldots,p^{\ell}\}) = \{p: Bp \geq \vec 0\}$ for $B \in \reals^{k \times n}$.
  We have $D(\T^r) = C$ if and only if both the halfspace and vertex conditions hold.
\end{theorem}
\begin{proof}
  ($\implies$)
  Suppose $D(\T^r) = C$.
  First, we note that the vertex condition is immediate: For all $j \in [\ell]$, $p^j \in C$ which gives $p^j \in D(\T^r)$.
  To show the halfspace condition is satisfied, we first construct a matrix $E$ such that $Ep \geq 0 \iff Bp \geq 0$, then use this construction to pick out the necessary vectors $v_1,\dots,v_k$.

  By Lemma \ref{lemma:minkowski-support}, there is a finite collection of vectors $w_1,\dots,w_{K} \in \reals^d$ and such that $\vec 0 \in \oplus_p \T^r$ if and only if, for all $w_i$, $\sum_y p(y) \max_{x \in T^r_y} \inprod{w_i}{x} \geq 0$.
  Hence, each vector $w_i$ generates a row of a matrix $E \in \reals^{K \times n}$ with $E_{iy} = \max_{x \in T^r_y} \inprod{w_i}{x}$, and we have $p \in D(\T^r) \iff Ep \geq 0$.
  By assumption of $D(\T^r) = C$, then, we have $Ep \geq 0 \iff Bp \geq 0$.
  By Lemma \ref{lemma:E-to-B}, because $B$ has the minimum possible number of rows, each row of $B$ appears (scaled by some positive constant) as a different row of $E$. Taking the collection of $w_i$ corresponding to these rows and rescaling them by that positive constant, we get a collection of $k$ vectors that we can rename $v_1,\ldots,v_k \in \reals^d$, with $\max_{x \in T^r_y} \inprod{v_i}{x} = B_{iy}$, hence the halfspace condition is satisfied.

  ($\impliedby$)
  Suppose Conditions \ref{cond:H-condition} and \ref{cond:V-condition} hold.
  Then by the vertex condition, $p^j \in D(\T^r)$ for all $j \in [\ell]$.
  Because $D(\T^r)$ is convex (Lemma \ref{lemma:D-polytope}), this implies $C \subseteq D(\T^r)$.
  To show $D(\T^r) \subseteq C$, let $p \in D(\T^r)$; by definition, $0 \in \oplus_p \T^r$.
  Then in particular for each vector $v_1,\ldots,v_k$ guaranteed by the halfspace condition, we have
  \begin{align*}
    0 &\leq \max_{x \in \oplus_p \T^r} \inprod{v}{x}  \\
      &=    \sum_{y \in \Y} p(y) \max_{x \in T^r_y} \inprod{v_i}{x}  \\
      &\leq \sum_{y \in \Y} p(y) B_{iy} .
  \end{align*}
  This proves $Bp \geq 0$, so $p \in C$.
\end{proof}

Given this characterization, we prove the main result on when a solution to the quadratic program exists.
\bo{Should note that the program's solution does give candidates $\T^r$, although their usefulness is unclear...}
\bo{Actually, I guess one could run this, then use monotonicity to add constraints and find the next subgradient sets, etc.}

\begin{proof}[Proof of Theorem \ref{thm:opt-iff-qfp}]
	\jessie{Move to appendix?}
    By Theorem \ref{thm:vertex-halfspace-opt}, it suffices to show that $\T^r$ satisfying the halfspace and vertex conditions exist if and only if the program is feasible.

  ($\implies$)
  By the vertex condition, for each $j \in [\ell]$, there exist witnesses $\{x^j_y \in T^r_y : y \in \Y\}$ satisfying the second constraint of the quadratic program (Inequality \ref{eqn:qp-constr-2}).
  By the halfspace condition, there exist normals $v_1, \dots, v_k$ such that, for all $i$, for all $x \in T^r_y$, $\inprod{v_i}{x} \leq B_{iy}$; in particular, this applies to the above witnesses $x^j_y \in T^r_y$.
  Collecting $v_1,\dots,v_k$ as the columns of $V$, this shows that the first constraint (Inequality \ref{eqn:qp-constr-1}) is satisfied.

>>>>

\paragraph{Embeddings.}
%We specifically study polyhedral functions in part because they have high-dimensional subgradient sets at points of nondifferentiability, which will be of interest to us later.
Polyhedral losses are motivated partly because they correspond to a natural surrogate construction technique: embedding the discrete predictions $\R$ as points in $\reals^d$ \jessie{and linearly interpolating between embedded reports}.
The key condition required on the embedding and surrogate loss $L$ is that a discrete prediction $r \in \R$ should minimize expected $\ell$-loss if and only if its embedding point minimizes expected $L$-loss.
%
\begin{definition}[Embedding of a loss]
  A loss $L: \reals^d \to \nonnegreals^{\Y}$ \emph{embeds} a loss $\ell: \R \to \nonnegreals^{\Y}$ \emph{in dimension $d$} with \emph{embedding function $\varphi: \R \to \reals^d$} if: (i) $\varphi$ is injective; (ii) for all $r \in \R$, $\ell(r) = L(\varphi(r))$; and (iii) for all $r \in \R$, $\gamma_r = \Gamma_{\varphi(r)}$.
  We simply say $L$ \emph{embeds} $\ell$ (in dimension $d$) if some such embedding function exists.
\end{definition}
%
Note that embedding does not immediately imply a key desirable property, consistency, which is equivalent to calibration in finite-outcome settings.
Informally, calibration means that if one minimizes the expected surrogate loss arbitrarily well over $\reals^d$, then applies the link function to obtain a discrete prediction $r$, then $r$ exactly minimizes $\ell$-loss.
\begin{definition}[Calibrated surrogate]\label{def:calibration}
  A surrogate loss $L: \reals^d \to \nonnegreals^{\Y}$ and \emph{link function} $\psi: \reals^d \to \R$ are \emph{calibrated} for a loss $\ell: \R \to \nonnegreals^{\Y}$ if for all $p \in \simplex$,
    \[ \inf_{u \in \reals^d ~:~ \psi(u) \not\in \gamma(p)} \inprod{p}{L(u)}  > \inf_{u \in \reals^d} \inprod{p}{L(u)}  .\]
  We simply say $L$ can be \emph{calibrated} to $\ell$ if such a link function exists.
\end{definition}
\bo{More on why calibration is desirable?}
\jessie{Taking a crack at this:}
As we alluded to above, calibration is not actually the trait we desire of a surrogate; rather, we seek consistency.
\cite{tewari2007consistency} show that, in a finite-outcome setting, the presence of calibration and consistency are equivalent conditions.
By consistency, we mean the standard notion where the expected loss over $\X \times \Y$ for a sequence of hypotheses $h_m$ converging to the optimal expected surrogate loss implies the expected discrete loss over the linked hypotheses $\psi \circ h_m$ converges to the optimal discrete loss.
The equivalence of these conditions allows us to focus on calibration, so we can abstract away the input space $\X$ and simply focus on probability distributions over the outcomes $\Y$.

@@

Therefore, we abstract to consider collections of sets $T^r_y$, which aspire to be the subgradient sets of a calibrated polyhedral surrogate $L(\cdot)_y$ at $u_r$.
Throughout, we fix $r$ and remove it from the subscript for ease of exposition.
Note that if $L(\cdot)_y$ is a polyhedral function on $\reals^d$, then all of its subgradient sets are (bounded) closed polytopes.\bo{cite}\jessie{\cite{rockafellar1997convex} ? in my notes, but I don't have the exact theorem number.}

\begin{definition}[$\T$, $D(\T)$]
  We use the following notation for a collection of closed polytopes, with implicit parameters $d$ and $r$: $\T = \{T_y \subseteq \reals^d : y \in \Y\}$.

  Given a distribution $p \in \simplex$, we write the $p$-weighted Minkowski sum of $\T$ as
    \[ \oplus_p \T := \left\{ \sum_{y \in \Y} p(y) x_y ~\Big|~ x_1 \in T_1, \dots, x_y \in T_y \right\} , \]
  in other words, the Minkowski sum of the scaled sets $\{p(y) T_y : y \in \Y\}$.

  Finally, we associate with $\T$ a set of distributions $D(\T) = \{ p \in \simplex : \vec 0 \in \oplus_p \T\}$.
\end{definition}

The importance of the $p$-weighted Minkowski sum and of $D(\T)$ are that they capture the distributions $p$ for which $u_r$ minimizes expected loss, assuming that $\T$ correspond to some polyhedral $L$.
\begin{proposition} \label{prop:minkowski-0-opt}
  If $T_y$ is the subgradient set of each $L(u_r)_y$, then $D(\T)$ is the set of distributions $p \in \simplex$ for which the embedding point $u_r$ minimizes expected $L$-loss.
\end{proposition}
\begin{proof}
  Any convex function $f$ is minimized at $u$ if and only if $\vec 0$ is a subgradient of $f$ at $u$.
  Let $f(u) = \inprod{p}{L(u)}$.
  Then the subgradient set of $f$ at $u$ is $\oplus_p \T$.
  This follows from the basic properties that if $f_1,f_2$ are convex with subgradient sets $T_1,T_2$ at $u$, then $\alpha f_1$ has subgradient set $\alpha T_1$ and $f_1 + f_2$ has subgradient set $T_1 \oplus T_2$, the Minkowski sum.
  So $\vec 0$ is a subgradient of $\inprod{p}{L(u_r)}$ if and only if $\vec 0 \in \oplus_p \T$, which by definition occurs if and only if $p \in D(\T)$.
\end{proof}
This fact will be vital for characterizing when $\ell$ is correctly embedded by some $L$ whose subgradient sets are $\T$ for each $r \in \R$.
We leverage it next.

@@

Two conditions are required: \emph{optimality} and \emph{monotonicity}.
Optimality enforces that the surrogate is minimized precisely when and where it should be.
It says that for each discrete prediction $r$ and set of distributions $\gamma_r$ for which it is optimal, there exists a collection of polytopes $\T$ (presumably the subgradient sets of $L(u_r)_y$ for each $y$) realizing Proposition~\ref{prop:minkowski-0-opt}.
Monotonicity says that these individual polytopes can actually be glued together to form the subgradients of some convex loss function $L$.

@@

\begin{theorem} \label{thm:general-char}
  Let $\ell: \R \to \reals_{\geq 0}^{\Y}$ be a discrete loss with, for each $r \in \R$, $\gamma_r = \{p : r \in \argmin_{r'} \inprod{p}{\ell(r)}\}$.
  Then $\ell$ is $d$-embeddable if and only if there exist polytopes $\T = \{T_y \subseteq \reals^d : y \in \Y\}$ for each $r \in \R$ such that both of the following hold:
  \begin{enumerate}
    \item (Optimality) We have $D(\T) = \gamma_r$.
    \item (Monotonicity) There exists an injective embedding function $\varphi : \R \to \reals^d$ and loss functions $\{L_y : \reals^d \to \reals^\Y_{\geq 0}\}_{y \in \Y}$ such that for all $y \in \Y$, we have $T_y = \partial L_y(\varphi(r))$.
  \end{enumerate}
\end{theorem}
\begin{proof}
  \jessie{An attempt:}
  $(\implies)$ If $\ell$ is $d$-embeddable, say that $L:\reals^d \to \reals^\Y_{\geq 0}$ embeds $\ell$.
  Take $T_y$ to be $\partial L(\varphi(r))_y$ for a fixed $r \in \R$ and $y \in \Y$.
  Clearly this set of polytopes $\T$ satisfies the monotonicity condition.
  To see the optimality condition is satisfied, consider that $\gamma_r = \Gamma_{\varphi(r)} = \{p \in \simplex : \vec 0 \in \oplus_p \partial L(\varphi(r))_y \} = \{p : \vec 0 \in \oplus_p T_y \} = D(\T)$.

  \bigskip
  $(\impliedby)$ The first embedding condition holds by the assumption of $\varphi$ in the monotonicity condition.
  Now consider that $\gamma_r = D(\T) = \{p \in \simplex : \vec 0 \in \oplus_p T_y \} = \{p \in \simplex : 0 \in \oplus_p \partial L(\varphi(r))_y \} = \{p : r \in \Gamma(p) \} =  \Gamma_{\varphi(r)}$.
  Therefore, as this holds for all $r \in \R$, we have that the Bayes Risks of the two losses match, and by~\cite[Proposition 1]{finocchiaro2019embedding}, we have $L$ embedding $\ell$. 
  \jessiet{This is a hole, since the subgradients of the level sets match, but the values do not.  Need to bake matching losses into monotonicity condition}
\end{proof}

@@

We now investigate the optimality condition of Theorem \ref{thm:general-char}, for two purposes.
First, we aim to greatly narrow the search space for constructing low-dimensional surrogate loss functions for a given discrete loss.
The tools we construct in this section aid in this task by constraining or constructing feasible subgradient sets $\T$ given a level set $\gamma_r$.
Second, we wish to prove impossibilities, i.e. lower bounds on the embedding dimension of a given discrete loss (an apparently hard problem).

By dropping monotonicity from Theorem \ref{thm:general-char}, we obtain $|\R|$ independent optimality conditions, $D(\T) = \gamma_r$, that must be satisfied in order to have a $d$-dimensional embedding.
For upper bounds, one must construct a $\T$ satisfying $D(\T) = \gamma_r$ for all $r$ (and then turn to satisfying monotonicity).
For lower bounds, it suffices to prove nonexistence of any such $\T$, for any $r$.

To accomplish these, we construct a quadratic feasibility program (QFP, Definition \ref{def:qfp}) that takes as input a polytope $C \subseteq \simplex$ and a parameter $d \in \mathbb{N}$ and (roughly) attempts to construct $\T$.
We will prove the following main results.
\begin{theorem} \label{thm:opt-iff-qfp}
  Given a convex polytope $C \subseteq \simplex$, there exist polytopes $\T$ in $\reals^d$ such that $D(\T) = C$ if and only if there is a feasible solution to the QFP (Definition \ref{def:qfp}) with parameter $d$.
\end{theorem}

@@

The variables are a set of normal vectors $\{v_i \in \reals^d : i \in [k]\}$ and a set of vertices $\{x^j_y \in \reals^d : j \in [\ell], y \in \Y\}$.
These correspond to a (possibly relaxed) halfspace representation and (possibly incomplete) vertex representation of the subgradient polytopes $\T$.

@@

To prove Theorem \ref{thm:opt-iff-qfp}, it is useful to develop a less-constructive intermediate characterization of when $D(\T) = C$.
The following condition formalizes a necessary condition on $\T$ in terms of the halfspace representation of $C$; the subsequent one formalizes a sufficient condition using the vertex representation.

\begin{condition}[Halfspace condition]\label{cond:H-condition}
	A collection of polytopes $\T$ and a polytope $C$ in the simplex defined by $C = \{p \in \simplex : Bp \geq \vec 0\}$ \emph{satisfy the halfspace condition} if there exist $v_1, \ldots, v_k \in \reals^d$ such that, for all $i \in [k]$ and $y \in \Y$, for all $x \in T_y$, we have $\inprod{v_i}{x} \leq B_{iy}$.
\end{condition}
\begin{condition}[Vertex condition]\label{cond:V-condition}
	A collection of polytopes $\T$ and a polytope $C$ in the simplex defined by $C = \conv(\{p^1, \ldots, p^\ell\})$ \emph{satisfy the vertex condition} if for all $j \in [\ell]$, $0 \in \oplus_{p^j} \T$. %% Equivalent: there exist $\{x_{jy} \in T_y : j \in [\ell]\}$ such that, for all $j \in [\ell]$, $\sum_{y \in \Y} p^j_y x_{jy} = \vec 0$.
\end{condition}

To gain a better understanding about the QFP in Definition~\ref{def:qfp}, we prove the following:
\begin{lemma} \label{lemma:D-polytope}
  For any $\T$, $D(\T)$ is a polytope (in particular, is convex).
\end{lemma}
\begin{proof}
   Recall by definition, the notation $\oplus_p \T = \{\sum_y p(y) x^r_y : x^r_y \in T_y (\forall y)\}$.
   Each $T_y$ is a polytope, so $p(y) T_y$ is a polytope.
   The Minkowski sum of polytopes is a polytope, so $\oplus_p \T$ is a polytope by~\cite[Section 1.2]{weibel2007minkowski}.
  \bo{TODO}
\end{proof}
\begin{lemma}  \label{lemma:minkowski-support}
  Given polytopes $\T$, there exists a finite set of normal vectors $w_1,\ldots,w_K \in \reals^d$ such that, for all $p \in \simplex$, $\oplus_p \T = \{x : \inprod{w_i}{x} \leq \sum_{y \in \Y} p(y) \max_{x \in T_y} \inprod{w_i}{x} \}$.
\end{lemma}
\begin{proof}
   For each $p$, $\oplus_p \T$ is a polytope.
  \bo{TODO: it turns out that $p$ doesn't matter, only its support; and there are only finitely many possible supports (or even better, the normals that are complete for the uniform minkowski sum turn out to be complete for all other $p$-combinations).}
\end{proof}
\begin{lemma} \label{lemma:E-to-B}
  Let $C = \{p : Bp \geq \vec 0 \}$ where $B$ has the minimum possible number of rows to capture $C$, and suppose $C = \{p : Ep \geq \vec 0 \}$.
  Then for each row in $B$ there is some (unique) row in $E$ that is equal to $\alpha B$ for some positive $\alpha$.
\end{lemma}
\begin{proof}
  \bo{? Cite} \jessie{Is left as an exercise in the following text:} \cite[Exercise 2.15]{ziegler2012lectures}
\end{proof}

\begin{theorem} \label{thm:vertex-halfspace-opt}
  Let the polytopes $\T = \{T_y \subseteq \reals^d : y \in \Y\}$ and $C$ be given, with $C = \conv(\{p^1,\ldots,p^{\ell}\}) = \{p: Bp \geq \vec 0\}$ for $B \in \reals^{k \times n}$.
  We have $D(\T) = C$ if and only if both the halfspace and vertex conditions hold.
\end{theorem}
\begin{proof}
  ($\implies$)
  Suppose $D(\T) = C$.
  First, we note that the vertex condition is immediate: For all $j \in [\ell]$, $p^j \in C$ which gives $p^j \in D(\T)$.
  To show the halfspace condition is satisfied, we first construct a matrix $E$ such that $Ep \geq 0 \iff Bp \geq 0$, then use this construction to pick out the necessary vectors $v_1,\dots,v_k$.

  By Lemma \ref{lemma:minkowski-support}, there is a finite collection of vectors $w_1,\dots,w_{K} \in \reals^d$ and such that $\vec 0 \in \oplus_p \T$ if and only if, for all $w_i$, $\sum_y p(y) \max_{x \in T_y} \inprod{w_i}{x} \geq 0$.
  Hence, each vector $w_i$ generates a row of a matrix $E \in \reals^{K \times n}$ with $E_{iy} = \max_{x \in T_y} \inprod{w_i}{x}$, and we have $p \in D(\T) \iff Ep \geq 0$.
  By assumption of $D(\T) = C$, then, we have $Ep \geq 0 \iff Bp \geq 0$.
  By Lemma \ref{lemma:E-to-B}, because $B$ has the minimum possible number of rows, each row of $B$ appears (scaled by some positive constant) as a different row of $E$. Taking the collection of $w_i$ corresponding to these rows and rescaling them by that positive constant, we get a collection of $k$ vectors that we can rename $v_1,\ldots,v_k \in \reals^d$, with $\max_{x \in T_y} \inprod{v_i}{x} = B_{iy}$, hence the halfspace condition is satisfied.

  ($\impliedby$)
  Suppose Conditions \ref{cond:H-condition} and \ref{cond:V-condition} hold.
  Then by the vertex condition, $p^j \in D(\T)$ for all $j \in [\ell]$.
  Because $D(\T)$ is convex (Lemma \ref{lemma:D-polytope}), this implies $C \subseteq D(\T)$.
  To show $D(\T) \subseteq C$, let $p \in D(\T)$; by definition, $0 \in \oplus_p \T$.
  Then in particular for each vector $v_1,\ldots,v_k$ guaranteed by the halfspace condition, we have
  \begin{align*}
    0 &\leq \max_{x \in \oplus_p \T} \inprod{v}{x}  \\
      &=    \sum_{y \in \Y} p(y) \max_{x \in T_y} \inprod{v_i}{x}  \\
      &\leq \sum_{y \in \Y} p(y) B_{iy} .
  \end{align*}
  This proves $Bp \geq 0$, so $p \in C$.
\end{proof}

Given this characterization, we prove the main result on when a solution to the quadratic program exists.
\bo{Should note that the program's solution does give candidates $\T$, although their usefulness is unclear...}
\bo{Actually, I guess one could run this, then use monotonicity to add constraints and find the next subgradient sets, etc.}

\begin{proof}[Proof of Theorem \ref{thm:opt-iff-qfp}]
	\jessie{Move to appendix?}
    By Theorem \ref{thm:vertex-halfspace-opt}, it suffices to show that $\T$ satisfying the halfspace and vertex conditions exist if and only if the program is feasible.

  ($\implies$)
  By the vertex condition, for each $j \in [\ell]$, there exist witnesses $\{x^j_y \in T_y : y \in \Y\}$ satisfying the second constraint of the quadratic program (Inequality \ref{eqn:qp-constr-2}).
  By the halfspace condition, there exist normals $v_1, \dots, v_k$ such that, for all $i$, for all $x \in T_y$, $\inprod{v_i}{x} \leq B_{iy}$; in particular, this applies to the above witnesses $x^j_y \in T_y$.
  Collecting $v_1,\dots,v_k$ as the columns of $V$, this shows that the first constraint (Inequality \ref{eqn:qp-constr-1}) is satisfied.

  \bigskip
  ($\impliedby$)
  We construct $T_y = \conv(\{x^1_y, \ldots, x^{\ell}_y\})$.
  The second constraint of the quadratic program immediately implies the vertex condition.
  Taking $v_1,\dots,v_k$ as the columns of $V$, the first constraint implies that for each $x^j_y$, we have $\inprod{v_i}{x^j_y} \leq B_{iy}$ for all $i,j,y$.
  Any point $x \in T_y$ is a convex combination of $x^1_y,\ldots,x^{\ell}_y$, so it satisfies $\inprod{v_i}{x} \leq B_{iy}$.
  This implies the halfspace condition.  % Bo: QED
\end{proof}
